---
title: "Analysis of high-throughput sequencing data from NIH GEO database"
author: "Taavi Päll^1^, Tanel Tenson^2^, Ülo Maiväli^2^"
date: '`r Sys.Date()`'
bibliography: references.bib
csl: jbc.csl
output:
  bookdown::html_book: 
    number_sections: FALSE
    toc: FALSE
    split_by: none
---

^1^ Department of Chemistry and Biotechnology, Tallinn University of Technology, Akadeemia tee 15, 12618 Tallinn, Estonia     
^2^ Institute of Technology, University of Tartu, Nooruse 1, 50411, Tartu, Estonia

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.path = 'figures/', dev = c('pdf'))
library(tidyverse)
library(lubridate)
library(stringr)
library(magrittr)
library(formattable)
library(Biobase)
library(grid)
library(gridExtra)
```

```{r libs}
# Load helper functions
source("lib/helpers.R")
# Set panel label case
panel_label_case <- "upper"
```

# Abstract

Assessing the quality of experiments, both directly and through replication attempts, is becoming a grave concern in biomedicine. 
This is especially true in omics experiments, where thousands of independent measurements are done in parallel, with the understanding that only a small minority will probe scientifically meaningful effects – a state of affairs conductive to mistaking false positives for scientific discoveries. 
A currently popular way to deal with false positives in omics is to calculate the Bayesian false-discovery rate (FDR) using, for the prior, the information present in the multitude of parallel measurements of a single omics experiment. 
To this picture we add a simple method to calculate a retrospective power that is compatible with the FDR. 
This allows to give a rich account of the quality and potential replicability of the results of almost any omics experiment. 
As proof of concept we assessed the quality of the __N__ high-throughput sequencing experiments published in Entrez GEO database.

